---
title: "Data Analysis"
author: "Natalie Tarn"
date: "2/12/2019"
output: github_document
---


### Motivation:
We want to test the perception of color by male ZBs, who use red coloration in dominance signaling. If perception is the same as females, it is fine and interesting in its own right. If different, it suggests that selection is acting differently on males and females who assess the same color variation. E.g., males assess dominance-related color variation differently from how females assess mate choice-related color variation.

## Goal 1 - test the dominance hierarchy of the males used in perception analyses
### Motivation/reasoning for Goal 1:

Although admittedly slightly unlikely, it could be that a male's perception of color variation is dependent upon his position in a hierarchy. 

Using the color categories we know females show (e.g., 1-5 v. 6-8), it could be that a male of high dominance status (and, potentially, redder coloration) would benefit from being able to assess fellow males that are closer to his own status and, presumably, more likely to displace him. In this way, males with coloration and dominance of a certain category may show better perception, or a different boundary, relative to their status/coloration. 

Broadly-speaking, perception could be related to an individual's own phenotype. In the case of males, the phenotype is both the male's own coloration and his dominance status. 

Importantly, this could happen even on a short-term basis, as some experimental work has shown that beak color and dominance status can change rapidly (i.e., within 3 days). 

### Steps: 
1) load dataset
2) look at variation in dominance interactions
3) look at overall network of dominance interactions
4) establish dominance rankings

```{r setup, include=FALSE}
d<-read.csv("Dominance_Hierarchy_Data_Test_Oct042018.csv")

```

### Variation in dominance interactions over time.

This plot shows how the number of interactions progresses along the timing of filming. Note that the number of interactions decreases each day of filming, i.e. the hierarchy likely stabilized.

```{r interaction-plot, message=FALSE,echo=FALSE}
plot(d$Date.int[d$Obs.Day==1]~d$time_sec[d$Obs.Day==1],type='b',xlim=c(0,1800),pch=19,
     xlab="time (s)",ylab="number of interaction")+
  points(d$Date.int[d$Obs.Day==2]~d$time_sec[d$Obs.Day==2],type='b',col='dark red',pch=19)+
  points(d$Date.int[d$Obs.Day==3]~d$time_sec[d$Obs.Day==3],type='b',col='dark blue',pch=19)
legend("topleft",c('Day 1','Day 2','Day 3'),pch=c(19,19,19),col=c('black','dark red','dark blue'))
```

### Network of dominance interactions. 

This plot shows a network of who is being aggressive toward whom. Larger circles represent indivdiuals that were involved in a greater number of interactions. Larger arrows represent a greater number of interactions. The legend shows which number corresponds to which individual

```{r dominance-network, echo=FALSE,warning=FALSE,message=FALSE}
library('sand')
library('igraph')
library('sna')
library('diagram')
sequence<-cbind(d$Aggressor,d$Receiver)
#head(sequence)
graph<-graph.data.frame(sequence,directed = T)
adjacency.matrix<-get.adjacency(graph,sparse=F)
#adjacency.matrix
network<-graph.adjacency(adjacency.matrix,weighted=TRUE,diag=FALSE)
plot.igraph(network,vertex.size=degree(adjacency.matrix)*1.5,edge.width=E(network)$weight/2, edge.arrow.size=ifelse(E(network)$weight==1,0.5,E(network)$weight*10),layout=layout_in_circle(network),edge.curved=TRUE)
legend('topleft',c("1=BLK","2=DKBLU","3=DKGRN","4=LTBLU","5=LTGRN","6=YELLOW"))
```

### Dominance hierarchy

This table shows the Elo rating for each individual - a measure of how dominant each individual is based on how often it was aggressive and how often it received aggression. 

```{r hierarchy, echo=FALSE,message=FALSE}
require(aniDom)
winners<-as.integer(d$Aggressor)
losers<-as.integer(d$Receiver)
scores<-as.data.frame(elo_scores(winners,losers))
rownames(scores)<-c('yellow','ltgrn','dkblu','dkgrn','blk','ltblu')
colnames(scores)<-"Elo rating"
sort(scores,decreasing = TRUE)
```

## Take-home messages
It looks like individual DKBLU was overall the most dominant, and BLK was the least dominant. We can come back to these ratings after seeing how these individuals do in perceptual trials, and ask if there is any correlation between dominance and perception. 

Importantly, DKBLU was the only bird to have a red band during our assessment - this is because he came from the Mooney Lab with a red band and we didn't remove the band before dominance assessment. I'd find it highly unlikely that this was the sole cause of his dominance and prior literature has found conflicting results as to whether band color influences dominance. However, this is something to consider (and, in future tests, we should remove red bands). 




