---
title: "Data Analysis"
author: "Natalie Tarn"
date: "2/12/2019"
output: html_document
---


### Motivation:
We want to test the perception of color by male ZBs, who use red coloration in dominance signaling. If perception is the same as females, it is fine and interesting in its own right. If different, it suggests that selection is acting differently on males and females who assess the same color variation. E.g., males assess dominance-related color variation differently from how females assess mate choice-related color variation.

## Goal 1 - test the dominance hierarchy of the males used in perception analyses
### Motivation/reasoning for Goal 1:

Although admittedly slightly unlikely, it could be that a male's perception of color variation is dependent upon his position in a hierarchy. 

Using the color categories we know females show (e.g., 1-5 v. 6-8), it could be that a male of high dominance status (and, potentially, redder coloration) would benefit from being able to assess fellow males that are closer to his own status and, presumably, more likely to displace him. In this way, males with coloration and dominance of a certain category may show better perception, or a different boundary, relative to their status/coloration. 

Broadly-speaking, perception could be related to an individual's own phenotype. In the case of males, the phenotype is both the male's own coloration and his dominance status. 

Importantly, this could happen even on a short-term basis, as some experimental work has shown that beak color and dominance status can change rapidly (i.e., within 3 days). 

### Steps: 
1) load dataset
2) look at variation in dominance interactions
3) look at overall network of dominance interactions
4) establish dominance rankings

```{r setup, include=FALSE}
library('sand')
library('igraph')
library('sna')
library('diagram')
library(tidyverse)
d<-read.csv("Dominance_Hierarchy_Data_Natalie.csv")

```

### Variation in dominance interactions over time.

This plot shows how the number of interactions progresses along the timing of filming. Note that the number of interactions decreases each day of filming, i.e. the hierarchy likely stabilized.

```{r, message=FALSE,echo=FALSE}
plot(d$Interaction_by_date[d$Obs.Day==1]~d$time_sec[d$Obs.Day==1],type='b',xlim=c(0,1800),pch=19,
     xlab="time (s)",ylab="number of interaction")+
  points(d$Interaction_by_date[d$Obs.Day==2]~d$time_sec[d$Obs.Day==2],type='b',col='dark red',pch=19)+
  points(d$Interaction_by_date[d$Obs.Day==3]~d$time_sec[d$Obs.Day==3],type='b',col='dark blue',pch=19)
legend("topleft",c('Day 1','Day 2','Day 3'),pch=c(19,19,19),col=c('black','dark red','dark blue'))
```

### Network of dominance interactions. 

This plot shows a network of who is being aggressive toward whom. Larger circles represent indivdiuals that were involved in a greater number of interactions. Larger arrows represent a greater number of interactions. The legend shows which number corresponds to which individual

```{r, echo=FALSE,warning=FALSE,message=FALSE}
sequence<-cbind(as.factor(d$Aggressor),as.factor(d$Receiver))
as.factor(d$Aggressor)
graph<-graph.data.frame(sequence,directed = T)
adjacency.matrix<-get.adjacency(graph,sparse=F)
rownames(adjacency.matrix)<-c("DBWT", "DBDB","WTDB","DBLB","DBLG","XXX","WTLG")
colnames(adjacency.matrix)<-c("DBWT", "DBDB","WTDB","DBLB","DBLG","XXX","WTLG")
network<-graph.adjacency(adjacency.matrix,weighted=TRUE,diag=FALSE)
plot.igraph(network,vertex.size=degree(adjacency.matrix)*1.5,edge.width=E(network)$weight/2, edge.arrow.size=ifelse(E(network)$weight==1,0.5,E(network)$weight*10),layout=layout_in_circle(network),edge.curved=TRUE)
legend('topleft',c("1=DBDB","2=DBLB","3=DBLG","4=DBWT","5=WTDB","6=WTLG", "7=XXX"))

```

### Dominance hierarchy

This table shows the Elo rating for each individual - a measure of how dominant each individual is based on how often it was aggressive and how often it received aggression. 

```{r, echo=FALSE,message=FALSE}
require(aniDom)
winners<-as.integer(d$Aggressor)
losers<-as.integer(d$Receiver)
scores<-as.data.frame(elo_scores(winners,losers))
rownames(scores)<-c('DBWT','DBDB','WTDB','DBLB','DBLG','XXX',"WTLG")
colnames(scores)<-"Elo rating"
sort(scores,decreasing = TRUE)
# Plot ranks
plot_ranks(scores,plot.CIs=TRUE)
# Plot hierachy shape
plot_hierarchy_shape(identity=1:nrow(scores),rank=1:nrow(scores),
winners=winners,losers=losers,fitted=TRUE)

```
## Other options in elo scores function
identities: Optional vector containing the identity of all individuals. This is useful if not all individuals are represented in the winners and losers, which we may want to use for WTLG, because it is not represented in the winners.
return.as.ranks: Boolean (TRUE/FALSE) describing whether to convert scores into ranks before
returning the data.
return.trajectories: Boolean (TRUE/FALSE) describing whether to return trajectories (the scores after each interaction T) or only the final scores
K: a parameter of the Elo function that determines the speed at which scores
change after an encounter (default K=200)
The elo score function can also be randomised (randomise=TRUE) to create K replicated datasets (where K=n.rands) to create a 95% confidence interval and improve estimates for ranks.

### Ardia et. al summary

## What predicts initial dominance position
Male zebra finches show consistency in aggression. Birds with greater red chroma initiated a higher number of aggressive interactions. Hue and body mass did not predict initial dominance position. Initial testosterone levels were also correlated with the number of interactions and red chroma. 

## How birds respond to a single exposure of Testosterone Propionate
Dominant birds (rank 1) injected with TP stayed dominant, but subordinate birds (rank 3) inreased their rank to rank 2 after the injection. TP-injected birds had increased red chroma in their bills for both dominant and subordinate birds. Hue increased in rank 3 birds. TP-injected birds had increased testosterone levels after 3 days, where rank 3 individuals had increased their testosterone levels more than rank 1 individuals. Birds that had higher levels of increased T were more aggressive, had a greater increase of red chroma, but not hue, UV hue, or brightness.

## Discussion - dynamic nature of bill color is mirrored by changes in testosterone levels (armament-ornament model)
In this study, we found high levels of red chroma and low levels of bill brightness, but not hue or UV hue, were correlated with dominance and aggression. Changes in bill color are driven partly by testosterone. Bill color plays a role (maybe as a signal or is correlated with other status reflecting traits) because the most aggressive and dominant birds had higher red chroma. An increased amount of testosterone leads to increased amounts of red chroma in bills, aggression, and changes in dominance rank. Male-male interactions in zebra finches may be more stable than previously thought, and there is potential for dominance hierarchies to exist.

However, bill color does not necessarily have a direct effect on dominance because they did not directly manipulate bill color.  We suggest to manipulate bill color directly to determine the direct role of bill color as a social signal. There is growing evidence that carotenoid-based signals can signal dominance.   

### Male bill colour and competition in zebra finches

## Abstract
Found no difference in behavior towards an orange compared to a red billed stimulus. However, the birds spent relatively more time eating when alone than compared to being close to a potential competitor. Additionally, more time was spent eating next to the orange/red billed stimulus as compared to the green billed stimulus.

## Results and Discussion
The presence of another male reduces the proportion of time spent eating. The red-orange difference may not be considered meaningful by the males.  They also probably ate less in front of the green-billed male because they of its abrnormality. This study supports the notion that beak color does not play a clear distinction in the results of the three experiments. Bill colour may play a role in male–male interactions, although the effect is limited and may be greater in the presence of females.

### Bill Color Preferences of Zebra Finches

## Abstract
Females tend to prefer males who have red, bright bills while males prefer females in the middle of the phenotypic range. Bill color appears to be more important in heterosexual than in isosexual interactions, and personal aesthetic preference seems to be the reason why certain mates are favored above others.

## Results and Discussion
Males have redder and darker beaks than females and females tended to associate with males with redder bills. When the beak of a male was painted artificially red to achieve the SR phenotype, females preferred that male.

### Take-home messages
It looks like individual DKBLU was overall the most dominant, ;and BLK was the least dominant. We can come back to these ratings after seeing how these individuals do in perceptual trials, and ask if there is any correlation between dominance and perception. 

Importantly, DKBLU was the only bird to have a red band during our assessment - this is because he came from the Mooney Lab with a red band and we didn't remove the band before dominance assessment. I'd find it highly unlikely that this was the sole cause of his dominance and prior literature has found conflicting results as to whether band color influences dominance. However, this is something to consider (and, in future tests, we should remove red bands). 




